---
title: "Hand crafting model matrix"
author: "Gan Yao"
date: "2022-12-15"
output: pdf_document
urlcolor: blue
---

This analysis attempts to manually create model matrix as suggested in [TR658](https://www.stat.umn.edu/geyer/aster/tr658.pdf) by Shaw et al. However, as it turns out, this is not necessary at least for this `echin2` data and our Lilium data, as we can use R formula to generate the exact same model matrix.

# Reproduction of `echin2` analysis in TR658

```{r message=FALSE}
library(aster)
library(tidyverse)
```

```{r}
data(echin2)
```

```{r}
vars <- c("lds1", "lds2", "lds3", "ld01", "ld02",
  "ld03", "roct2003", "ld04", "roct2004", "ld05",
  "roct2005")
pred <- c(0, 1, 2, 3, 4, 5, 6, 6, 8, 8, 10)
fam <- c(1, 1, 1, 1, 1, 1, 3, 1, 3, 1, 3)
nind <- length(unique(echin2$id))
nnode <- length(levels(echin2$varb))
```

```{r}
x <- echin2$resp
dim(x) <- c(nind, nnode)
r <- 0 * x + 1
```

## Hand-crafted model matrix

```{r}
modmat.super <- NULL
names.super <- NULL
for (i in levels(echin2$varb)) {
 modmat.super <- cbind(modmat.super, as.numeric(echin2$varb ==i))
 names.super <- c(names.super, i)
 }
```

```{r}
in.greenhouse <- is.element(echin2$varb, grep("lds", levels(echin2$varb), value = TRUE))
print(unique(echin2$varb[in.greenhouse]), max.levels = 0)
print(unique(echin2$varb[!in.greenhouse]), max.levels = 0)
```

```{r}
for (i in levels(echin2$flat)) if (i > "1") {
 modmat.super <- cbind(modmat.super, as.numeric(in.greenhouse &
 echin2$flat == i))
 names.super <- c(names.super, paste("flat", i,
 sep = ""))
 }
```


```{r}
for (i in levels(echin2$row)) if (i > "10") {
 modmat.super <- cbind(modmat.super, as.numeric((!in.greenhouse) &
 echin2$row == i))
 names.super <- c(names.super, paste("row", i,
 sep = ""))
 }
```

```{r}
for (i in levels(echin2$yearcross)) if (i >= "2000") {
 modmat.super <- cbind(modmat.super, as.numeric(echin2$yearcross ==i))
 names.super <- c(names.super, paste("yc", i, sep = ""))
 }
```

```{r}
modmat.super <- cbind(modmat.super, as.numeric(!in.greenhouse) *  echin2$posi)
names.super <- c(names.super, "posi")
```

```{r}
for (i in levels(echin2$crosstype)) if (i > "W") {
 modmat.super <- cbind(modmat.super, as.numeric(echin2$crosstype == i & echin2$varb == "roct2005"))
 names.super <- c(names.super, paste("cross", i, sep = ""))
 }
```

```{r}
for (i in levels(echin2$crosstype)) if (i > "W") {
 modmat.super <- cbind(modmat.super, as.numeric(echin2$crosstype == i & echin2$varb == "lds3"))
 names.super <- c(names.super, paste("crossgreen", i, sep = ""))
 }
```

```{r}
nodename <- unique(as.character(echin2$var))
modmat.super <- array(as.vector(modmat.super), c(dim(x), length(names.super)))
dimnames(modmat.super) <- list(NULL, nodename, names.super)
```

## Model fitted with handcrafted model matrix


```{r}
out.super <- aster(x, r, pred, fam, modmat.super)
summary(out.super)
```



# Model matrix by formula



```{r eval=FALSE, include=FALSE}
echin2$inGreenHouse <- as.numeric(in.greenhouse)
echin2$outGreenHouse <- as.numeric(!in.greenhouse)
echin2$fitField <- as.numeric(echin2$varb == 'roct2005')
echin2$fitChamber <- as.numeric(echin2$varb == 'lds3')
out.super.test <- aster(resp ~ -1 + varb + yearcross +inGreenHouse:flat 
                        + outGreenHouse:row + outGreenHouse:posi 
                        + fitField:crosstype + fitChamber:crosstype,
                        pred, fam, varb, id, root, data=echin2) 
```

```{r eval=FALSE, include=FALSE}
summary(out.super.test, info.tol=1e-12)
```

By creating necessary variables and applying R formula, we can get the exact same model.

```{r}
resp <- echin2$resp
varb <- echin2$varb
inGreenHouse <- as.numeric(in.greenhouse)
outGreenHouse <- as.numeric(!in.greenhouse)
fitField <- as.numeric(echin2$varb == 'roct2005')
fitChamber <- as.numeric(echin2$varb == 'lds3')
yearcross <- echin2$yearcross
flat2 <- as.numeric(echin2$flat == 2)
flat3 <- as.numeric(echin2$flat == 3)
row11 <- as.numeric(echin2$row == 11)
row12 <- as.numeric(echin2$row == 12)
row13 <- as.numeric(echin2$row == 13)
crossWi <- as.numeric(echin2$crosstype == 'Wi')
crossWr <- as.numeric(echin2$crosstype == 'Wr')
posi <- echin2$posi
```

```{r}
out.super.test_2 <- aster(resp ~ -1 + varb + yearcross +inGreenHouse:flat2 + inGreenHouse:flat3 
                        + outGreenHouse:row11 + outGreenHouse:row12 + outGreenHouse:row13 + outGreenHouse:posi 
                        + fitField:crossWi + fitField:crossWr + fitChamber:crossWi + fitChamber:crossWr,
                        pred, fam, varb, echin2$id, echin2$root)
```

```{r}
sum(is.na(echin2))
```

```{r}
summary(out.super.test_2)
```

`out.super` and `out.super.test_2` are the same.






