---
title: "Model Selection"
author: "Gan Yao"
date: "2023-01-20"
output: pdf_document
---

This analysis conducts model selection by AIC on models that are of our particular interest.


# Aaster graph

$$root \rightarrow flCt \rightarrow flCtNotConsumed \rightarrow flCtUndamaged \rightarrow capsuleCt \rightarrow isHarvested \rightarrow ovuleCt \rightarrow embryoCt(fitness)$$

# Load library and data, transfer data

```{r message=FALSE}
library(aster)
library(tidyverse)
```

```{r}
data <- read.csv("data/output/remLilium2021Data30Nov2022.csv")
names(data)
```

```{r}
data <- data[data$site != "lf",]
data <- data[data$site != "wrrx",]
data[is.na(data$nCapsulesHarvested), 'nCapsulesHarvested'] <- 0
data[is.na(data$ovuleCt), 'ovuleCt'] <- 0
data[is.na(data$embryoCt), 'embryoCt'] <- 0
names(data)[names(data) == 'nCapsulesHarvested'] <- 'isHarvested'
```

```{r}
pred <- c(0,1,2,3,4,5,6)
fam <- c(2,1,1,1,1,2,1)
vars <- c("flCt", "flCtNotConsumed","flCtUndamaged", "capsuleCt", 
          "isHarvested", "ovuleCt", "embryoCt")
test <- data %>% mutate(nn5Dist_s = nn5Dist/1000, 
                        nn5DistNotConsumed =replace_na(nn5DistNotConsumed, 0)) %>% 
  mutate(nn5DistNotConsumed_s = nn5DistNotConsumed/1000)

redata <- reshape(test, varying = list(vars), direction="long", timevar="varb", 
                  times = as.factor(vars), v.names="resp")

redata <- data.frame(redata, root = 1)
redata$fit <- as.numeric(redata$varb == "embryoCt")
redata$Nid <- as.numeric(gsub("[^0-9.-]", "", redata$id))

names(redata)
```

```{r}
redata$Deer <- as.numeric(redata$varb=='flCtNotConsumed')
redata$Pollination <- as.numeric(is.element(redata$varb, 
                                c("capsuleCt", "isHarvested", "ovuleCt", "embryoCt")))

names(redata)
```

# Null Model

```{r}
model.null <- aster(resp ~ -1 + varb, 
                pred, fam,varb,id,root,data=redata)
#summary(model3, info.tol=1e-12)
summary(model.null)

eigen(model.null$fisher)$val
```

# Provious Models

```{r}
model.P1 <- aster(resp ~ -1 + varb + fit:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

summary(model.P1)
```

```{r}
model.P2 <- aster(resp ~ -1 + varb + fit:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)

summary(model.P2)
```

```{r}
model.P3 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + fit:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)

summary(model.P3)
```

# Model1: $Deer \sim 1$, $Pollination \sim nnA$

```{r}
model1 <- aster(resp ~ -1 + varb + Pollination:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)
summary(model1)

eigen(model1$fisher)$val
```

# Model2: $Deer \sim 1$ and $Pollination \sim nnB$

```{r}
model2 <- aster(resp ~ -1 + varb + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)
summary(model2)

eigen(model2$fisher)$val
```

# Model3: $Deer \sim nnA$ and $Pollination \sim 1$

```{r}
model3 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)
summary(model3)

eigen(model3$fisher)$val
```

# Model4: $Deer \sim nnA$ and $Pollination \sim nnA$

```{r}
model4 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s + Pollination:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)
summary(model4, info.tol=1e-9)

eigen(model4$fisher)$val
```

# Model5: $Deer \sim nnA$ and $Pollination \sim nnB$

```{r}
model5 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)
summary(model5)

eigen(model5$fisher)$val
```

# Model Selection

Wrap up a function for computing AIC of aster models.

```{r}
aster_AIC <- function(mod) {
  return(mod$deviance + 2*length(mod$coefficients))
}
```

```{r echo=FALSE}
print(model.null$formula)
aster_AIC(model.null)


print(model.P1$formula)
aster_AIC(model.P1)


print(model.P2$formula)
aster_AIC(model.P2)


print(model.P3$formula)
aster_AIC(model.P3)


print(model1$formula)
aster_AIC(model1)


print(model2$formula)
aster_AIC(model2)


print(model3$formula)
aster_AIC(model3)


print(model4$formula)
aster_AIC(model4)


print(model5$formula)
aster_AIC(model5)
```

\
\
\
\
\
\
\
\
\
\

+------------+---------------------------------------------------------------+----------------+-------------------------------+
| Model      | Formula                                                       | AIC            | P-values of interaction terms |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 1          | $$                                                            | -249884.7\     | NA                            |
|            |                   resp \sim varb                              |                |                               |
|            |                   $$                                          |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 2          | $$                                                            | -249890.7\     | 0.7e-3                        |
|            |                   resp \sim varb + Pollination:nnA            |                |                               |
|            |                   $$                                          |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 3          | $$                                                            | **-249906.9**\ | 1.83e-9                       |
|            |                   resp \sim varb + Pollination:nnB            |                |                               |
|            |                   $$                                          |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 4          | $$                                                            | -249890.1\     | 0.00298                       |
|            |                   resp \sim varb + Deer:nnA                   |                |                               |
|            |                   $$                                          |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 5          | $$                                                            | -249891.1\     | 0.0997, 0.0667                |
|            |                   resp \sim varb + Deer:nnA + Pollination:nnA |                |                               |
|            |                   $$                                          |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 6          | $$                                                            | **-249905.8**\ | 0.341, 1.8e-6                 |
|            |          resp \sim varb + Deer:nnA + Pollination:nnB          |                |                               |
|            |          $$                                                   |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 7          | $$                                                            | -249886.9\     | 0.0186                        |
|            | resp \sim varb + fit:nnA                                      |                |                               |
|            | $$                                                            |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 8          | $$                                                            | **-249899.3**\ | 1.09e-6                       |
|            | resp \sim varb + fit:nnB                                      |                |                               |
|            | $$                                                            |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+
| 9          | $$                                                            | -249898.2\     | 0.334, 3.71e-6                |
|            | resp \sim varb + fit:nnA + fit:nnB                            |                |                               |
|            | $$                                                            |                |                               |
+------------+---------------------------------------------------------------+----------------+-------------------------------+

# 
