---
title: "Mode models and landscape"
author: "Gan Yao"
date: "`r Sys.Date()`"
output: pdf_document
---

This analysis include more models that are of our interest, as well as generating fitness landscape of these models.

# Aaster graph

$$root \rightarrow flCt \rightarrow flCtNotConsumed \rightarrow flCtUndamaged \rightarrow capsuleCt \rightarrow isHarvested \rightarrow ovuleCt \rightarrow embryoCt(fitness)$$

# Load library and data, transfer data

```{r message=FALSE, warning=FALSE}
library(aster)
library(tidyverse)
library(plotly)
library(hrbrthemes)
library(viridis)
library(gridExtra)
library(metR)
```

```{r}
data <- read.csv("data/output/remLilium2021Data30Nov2022.csv")
names(data)
```

```{r}
data <- data[data$site != "lf",]
data <- data[data$site != "wrrx",]
data[is.na(data$nCapsulesHarvested), 'nCapsulesHarvested'] <- 0
data[is.na(data$ovuleCt), 'ovuleCt'] <- 0
data[is.na(data$embryoCt), 'embryoCt'] <- 0
names(data)[names(data) == 'nCapsulesHarvested'] <- 'isHarvested'
```

```{r}
pred <- c(0,1,2,3,4,5,6)
fam <- c(2,1,1,1,1,2,1)
vars <- c("flCt", "flCtNotConsumed","flCtUndamaged", "capsuleCt", 
          "isHarvested", "ovuleCt", "embryoCt")
#test <- data %>% mutate(nn5Dist_s = nn5Dist/1000, 
                        #nn5DistNotConsumed =replace_na(nn5DistNotConsumed, 0)) %>% 
  #mutate(nn5DistNotConsumed_s = nn5DistNotConsumed/1000)

test <- data %>% mutate(nn5Dist_s = log(nn5Dist)/10, 
                        nn5DistNotConsumed =replace_na(nn5DistNotConsumed, 1)) %>% 
  mutate(nn5DistNotConsumed_s = log(nn5DistNotConsumed)/10)

redata <- reshape(test, varying = list(vars), direction="long", timevar="varb", 
                  times = as.factor(vars), v.names="resp")

redata <- data.frame(redata, root = 1)
redata$fit <- as.numeric(redata$varb == "embryoCt")
redata$Nid <- as.numeric(gsub("[^0-9.-]", "", redata$id))

names(redata)
```

```{r}
redata$Deer <- as.numeric(redata$varb=='flCtNotConsumed')
redata$Pollination <- as.numeric(is.element(redata$varb, 
                                c("capsuleCt", "isHarvested", "ovuleCt", "embryoCt")))

names(redata)
```

```{r}
# extreme_nnA <- unique(redata[redata$nn5Dist_s < 0,'id'])
# redata <- redata[(!redata$id %in% extreme_nnA), ]
```



# Models

## Model 0: Null Model

```{r}
model.null <- aster(resp ~ -1 + varb, 
                pred, fam,varb,id,root,data=redata)
#summary(model3, info.tol=1e-12)
summary(model.null)

eigen(model.null$fisher)$val
```

## Model 1: $Fitness:nnA$

```{r}
model1 <- aster(resp ~ -1 + varb + fit:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

summary(model1)
```

## Model 2: $Deer:nnA$

```{r}
model2 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

summary(model2)
```

## Model 3: $Pollination:nnA$

```{r}
model3 <- aster(resp ~ -1 + varb + Pollination:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

#summary(model3, info.tol=1e-9)
summary(model3)
```

## Model 4: $Fitness:nnA + Deer:nnA$

```{r}
model4 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

summary(model4)
```

## Model 5: $Fitness:nnA + Pollination:nnA$

```{r}
model5 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Pollination:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

#summary(model5, info.tol=1e-9)
summary(model5)
```

## Model 6: $Deer:nnA + Pollination:nnA$

```{r}
model6 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s + Pollination:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

#summary(model6, info.tol=1e-9)
summary(model6)
```

## Model 7: $Fitness:nnA + Deer:nnA + Pollination:nnA$

```{r}
model7 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s, 
                pred, fam,varb,id,root,data=redata)

#summary(model7, info.tol=1e-9)
summary(model7)
```

## Model 8: $Pollination:nnB$

```{r}
model8 <- aster(resp ~ -1 + varb + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)

summary(model8)
```

## Model 9: $Fitness:nnA + Pollination:nnB$

```{r}
model9 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)

summary(model9)
```

## Model 10: $Deer:nnA + Pollination:nnB$

```{r}
model10 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)

summary(model10)
```

## Model 11: $Fitness:nnA + Deer:nnA + Pollination:nnB$

```{r}
model11 <- aster(resp ~ -1 + varb + fit:nn5Dist_s +Deer:nn5Dist_s + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata)

summary(model11)
```

# Model Selection

```{r}
aster_AIC <- function(mod) {
  return(mod$deviance + 2*length(mod$coefficients))
}
```

```{r echo=FALSE}
print(model.null$formula)
aster_AIC(model.null)


print(model1$formula)
aster_AIC(model1)


print(model2$formula)
aster_AIC(model2)


print(model3$formula)
aster_AIC(model3)


print(model4$formula)
aster_AIC(model4)


print(model5$formula)
aster_AIC(model5)


print(model6$formula)
aster_AIC(model6)


print(model7$formula)
aster_AIC(model7)


print(model8$formula)
aster_AIC(model8)


print(model9$formula)
aster_AIC(model9)


print(model10$formula)
aster_AIC(model10)


print(model11$formula)
aster_AIC(model11)
```

# 1-D Fitness landscape



```{r}
fitness_landscape <- function(model, covariate = 'nn5Dist_s', lower = NULL, upper = NULL, 
                              observation=FALSE, scale_back = FALSE) {
  # Make fake individuals
  nInd <- 50
  lwr <- if (is.null(lower)) min(unique(model$data[,covariate])) else lower
  upr <- if (is.null(upper)) max(unique(model$data[,covariate])) else upper
  cand.nnA <- seq(from = lwr, to = upr, length = nInd)
  cand <- as.data.frame(cand.nnA)
  colnames(cand) <- covariate
  cand$root <- 1
  blah <- data[1:nInd, colnames(data) %in% vars]
  cand <- cbind(cand, blah)
  cand$id <- data[1:nInd, 'id']
  
  # Transform fake data into long format
  cand_long <- reshape(cand, varying = list(vars), direction="long", timevar="varb", 
                  times = as.factor(vars), v.names="resp")

  cand_long <- data.frame(cand_long)
  cand_long$fit <- as.numeric(cand_long$varb == "embryoCt")
  cand_long$Nid <- as.numeric(gsub("[^0-9.-]", "", cand_long$id))
  
  cand_long$Deer <- as.numeric(cand_long$varb=='flCtNotConsumed') 
  cand_long$Pollination <- as.numeric(is.element(cand_long$varb, 
                                c("capsuleCt", "isHarvested", "ovuleCt", "embryoCt")))
  #cand_long$Rtail <- as.numeric(cand_long$nn5Dist_s > quantile(cand_long$nn5Dist_s, 0.975)) 
  #cand_long$Ltail <- as.numeric(cand_long$nn5Dist_s < quantile(cand_long$nn5Dist_s, 0.025)) 
  
  # Get conditional mean value parameters
  pred <- predict(model, cand_long, varvar=varb, idvar=id, root=root,
                      se.fit = TRUE, model.type='conditional',
                  is.always.parameter = TRUE, info.tol=1e-8)
  
  xi_parm <- pred$fit
  xi_parm_se <- pred$se.fit
  
  names(xi_parm) <- paste0(cand_long$id,'.',cand_long$varb)
  names(xi_parm_se) <- paste0(cand_long$id,'.',cand_long$varb)
  
  xi_parm_grad <- pred$gradient
  rownames(xi_parm_grad) <- paste0(cand_long$id,'.',cand_long$varb)
  colnames(xi_parm_grad) <- names(model$coefficients)
  
  # Expected fitness
  Ids <- unique(cand_long$id)
  exp_fitness <- rep(0,nInd)
  names(exp_fitness) <- Ids
  for (id in Ids) {
    exp_fitness[id] <- prod(xi_parm[grep(paste0(id,'\\.'), names(xi_parm))][-5])
  }
  
  # Get covariance matrix of xi
  xi_parm_var = xi_parm_grad %*% solve(model$fisher) %*% t(xi_parm_grad)

  # Delta Method
  var_expfit <- rep(0,nInd)
  names(var_expfit) <- Ids
  for (id in Ids) {
    ind <- paste0(id,'\\.')
    xi <- xi_parm[grep(ind, names(xi_parm))][-5]
    var <- xi_parm_var[grep(ind, rownames(xi_parm_var))[-5], 
                        grep(ind, colnames(xi_parm_var))[-5]]
    grad <- c(prod(xi[-1]), prod(xi[-2]), prod(xi[-3]), prod(xi[-4]), prod(xi[-5]), prod(xi[-6]))
    var_expfit[id] <- t(grad) %*% var %*% grad
  }
  
  se_expfit = sqrt(var_expfit)
  xlabel = if (covariate == 'nn5Dist_s') 'log(nnA)/10' else 'log(nnB)/10'
  cand_block <- cand %>% mutate(exp_fitness, exp_fitness, lower = exp_fitness - 2 * se_expfit, upper = exp_fitness + 2 * se_expfit)
  if (scale_back == TRUE) {
    cand_block[, as.character(covariate)] <- exp(10*cand_block[,as.character(covariate)])
    xlabel = if (covariate == 'nn5Dist_s') 'nnA' else 'nnB'
  }
  #cand_block <- cand_block[1:40,]
  plt <- ggplot(data = cand_block) + geom_line(mapping = aes(x = cand_block[,as.character(covariate)], y = exp_fitness)) + geom_ribbon(mapping = aes(x = cand_block[,as.character(covariate)], ymin = lower, ymax = upper), alpha = 0.5) 
  if (observation == TRUE) {
    obs <- test %>% filter((!!sym(covariate)) > lwr & (!!sym(covariate)) < upr)
    if (scale_back == TRUE) {
      plt <- plt + geom_point(data=obs, mapping = aes(x =exp(10*obs[,as.character(covariate)]), 
                                                      y = obs[,'fecundity']))
    } else {
      plt <- plt + geom_point(data=obs, mapping = aes(x =obs[,as.character(covariate)], y = obs[,'fecundity']))
    }
  }
  plt <- plt + labs(x=xlabel, y="Expected fitness", title = paste0("Fitness landscape    ", substr(format(model$formula), start=20, stop=nchar(format(model$formula)))))  + 
    #annotate(geom='text', x=c(0, 0.1, 0.3, 0.5), y=750, label=c('nnA', '2.718282', '20.085537', '148.413159')) +
    theme_minimal() + scale_x_continuous(minor_breaks = NULL) +scale_y_continuous(minor_breaks = NULL)
  
  print(plt)
  
}
```

```{r}
#par(mar = c(4, 4, .1, .1))
fitness_landscape(model1, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model2, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model3, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model4, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model5, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model6, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model7, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model8, covariate = 'nn5DistNotConsumed_s', lower = quantile(test$nn5DistNotConsumed_s, 0.025), upper = quantile(test$nn5DistNotConsumed_s, 0.975))
```

```{r}
#par(mar = c(4, 4, .1, .1))
fitness_landscape(model1, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model2, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model3, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model4, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model5, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model6, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model7, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE)
fitness_landscape(model8, covariate = 'nn5DistNotConsumed_s', lower = quantile(test$nn5DistNotConsumed_s, 0.025), upper = quantile(test$nn5DistNotConsumed_s, 0.975), observation = TRUE)
```


```{r}
#par(mar = c(4, 4, .1, .1))
fitness_landscape(model1, observation = TRUE, scale_back = TRUE)
fitness_landscape(model2, observation = TRUE, scale_back = TRUE)
fitness_landscape(model3, observation = TRUE, scale_back = TRUE)
fitness_landscape(model4, observation = TRUE, scale_back = TRUE)
fitness_landscape(model5, observation = TRUE, scale_back = TRUE)
fitness_landscape(model6, observation = TRUE, scale_back = TRUE)
fitness_landscape(model7, observation = TRUE, scale_back = TRUE)
fitness_landscape(model8, covariate = 'nn5DistNotConsumed_s', observation = TRUE, scale_back = TRUE)
```



# 2-D Fitness landscape

```{r}
fitness_landscape_2d <- function(model) {
  # Make fake individuals
  nInd <- 50
  lwr_1 <-quantile(log(data$nn5Dist)/10, 0.025)
  upr_1 <-quantile(log(data$nn5Dist)/10, 0.975)
  lwr_2 <-quantile(test$nn5DistNotConsumed_s, 0.025)
  upr_2 <-quantile(test$nn5DistNotConsumed_s, 0.975)
  cand.nnA <- seq(from = lwr_1, to = upr_1, length = nInd)
  cand.nnB <- seq(from = lwr_2, to = upr_2, length = nInd)
  cand <- expand.grid(cand.nnA, cand.nnB)
  colnames(cand) <- c('nn5Dist_s', 'nn5DistNotConsumed_s')
  cand <- as.data.frame(cand)
  cand$root <- 1
  blah <- data[1:nInd, colnames(data) %in% vars]
  blah <- do.call("rbind", replicate(nInd, blah, simplify = FALSE))
  cand <- cbind(cand, blah)

  cand$id <- paste0('LP', seq(nInd^2))
  
  # Transform fake data into long format
  cand_long <- reshape(cand, varying = list(vars), direction="long", timevar="varb", 
                  times = as.factor(vars), v.names="resp")

  cand_long <- data.frame(cand_long)
  cand_long$fit <- as.numeric(cand_long$varb == "embryoCt")
  cand_long$Nid <- as.numeric(gsub("[^0-9.-]", "", cand_long$id))
  
  cand_long$Deer <- as.numeric(cand_long$varb=='flCtNotConsumed')
  cand_long$Pollination <- as.numeric(is.element(cand_long$varb, 
                                c("capsuleCt", "isHarvested", "ovuleCt", "embryoCt")))
  
  # Get conditional mean value parameters
  pred <- predict(model, cand_long, varvar=varb, idvar=id, root=root,
                      se.fit = TRUE, model.type='conditional',
                  is.always.parameter = TRUE, info.tol=1e-8)
  
  xi_parm <- pred$fit
  xi_parm_se <- pred$se.fit
  
  names(xi_parm) <- paste0(cand_long$id,'.',cand_long$varb)
  names(xi_parm_se) <- paste0(cand_long$id,'.',cand_long$varb)
  
  xi_parm_grad <- pred$gradient
  rownames(xi_parm_grad) <- paste0(cand_long$id,'.',cand_long$varb)
  colnames(xi_parm_grad) <- names(model$coefficients)
  
  # Expected fitness
  Ids <- unique(cand_long$id)
  exp_fitness <- rep(0,nInd^2)
  names(exp_fitness) <- Ids
  for (id in Ids) {
    exp_fitness[id] <- prod(xi_parm[grep(paste0(id,'\\.'), names(xi_parm))][-5])
  }
  
  # Get covariance matrix of xi
  xi_parm_var = xi_parm_grad %*% solve(model$fisher) %*% t(xi_parm_grad)

  # Delta Method
  var_expfit <- rep(0,nInd^2)
  names(var_expfit) <- Ids
  for (id in Ids) {
    ind <- paste0(id,'\\.')
    xi <- xi_parm[grep(ind, names(xi_parm))][-5]
    var <- xi_parm_var[grep(ind, rownames(xi_parm_var))[-5], 
                        grep(ind, colnames(xi_parm_var))[-5]]
    grad <- c(prod(xi[-1]), prod(xi[-2]), prod(xi[-3]), prod(xi[-4]), prod(xi[-5]), prod(xi[-6]))
    var_expfit[id] <- t(grad) %*% var %*% grad
  }
  
  se_expfit = sqrt(var_expfit)
  
  # Plots
  cand_block <- cand %>% mutate(exp_fitness, exp_fitness, lower = exp_fitness - se_expfit, upper = exp_fitness + se_expfit)
  
  
  # p1 <- ggplot(data=cand_block, aes(nn5Dist_s, nn5DistNotConsumed_s, fill=lower)) + 
  # geom_tile() +ggtitle(paste0("Fitness landscape    ", substr(format(model$formula), start=20, stop=nchar(format(model$formula)))))  +
  # scale_fill_gradient(low="white", high="blue") + 
  # theme(plot.title = element_text(size=10))
  
  # p2 <- ggplot(data=cand_block, aes(nn5Dist_s, nn5DistNotConsumed_s, fill=upper)) + 
  # geom_tile() +ggtitle(paste0("Fitness landscape    ", substr(format(model$formula), start=20, stop=nchar(format(model$formula)))))  +
  # scale_fill_gradient(low="white", high="blue") + 
  #  theme(plot.title = element_text(size=10))
  
  # p3 <- ggplot(data=cand_block, aes(nn5Dist_s, nn5DistNotConsumed_s, fill=exp_fitness)) +
  # geom_tile() +ggtitle(paste0("Fitness landscape    ", substr(format(model$formula), start=20, stop=nchar(format(model$formula)))))  +
  # scale_fill_gradient(low="white", high="blue") +
  #  theme(plot.title = element_text(size=10))
  
  p3 <- ggplot(cand_block, aes(nn5Dist_s, nn5DistNotConsumed_s, z= exp_fitness, colour=stat(level))) +
          geom_contour() +
          geom_text_contour(aes(z = exp_fitness)) +
          scale_colour_distiller(palette = 'Spectral', direction = 1) +
          scale_x_continuous(limits = c(lwr_1, upr_1))
  print(p3)
  
  
  
  #return(list(lwr = p1, upr = p2, exp = p3))
  
  #ggplotly(p3)
  
}
```


## Fitness landscape of model 9
```{r warning=FALSE}
res <- fitness_landscape_2d(model9)
# res$lwr
# res$exp
# res$upr
```

## Fitness landscape of model 10
```{r warning=FALSE}
res <- fitness_landscape_2d(model10)

#res$lwr
#res$exp
#res$upr
#grid.arrange(res$lwr ,res$exp ,res$upr, ncol=2, nrow=2)
```


## Fitness landscape of model 11
```{r warning=FALSE}
res <- fitness_landscape_2d(model11)
# res$lwr
# res$exp
# res$upr
```

# Handling extreme values

We observed abnormal behaviors on fitness landscapes of `model7`, our best model chosen by AIC,  at both right and left tails. So we try to handle these extreme values of `nn5Dist` study how they affect model performance.

```{r}
hist(log(data$nn5Dist)/10, breaks=20)
```

```{r}
wide2long <- function(org_data) {
  pred <- c(0,1,2,3,4,5,6)
  fam <- c(2,1,1,1,1,2,1)
  vars <- c("flCt", "flCtNotConsumed","flCtUndamaged", "capsuleCt", 
            "isHarvested", "ovuleCt", "embryoCt")
  
  test <- org_data %>% mutate(nn5Dist_s = log(nn5Dist)/10, 
                          nn5DistNotConsumed =replace_na(nn5DistNotConsumed, 1)) %>% 
    mutate(nn5DistNotConsumed_s = log(nn5DistNotConsumed)/10) %>% 
    mutate(Ltail = ifelse(nn5Dist_s < quantile(nn5Dist_s, 0.025), 1, 0),
                        Rtail = ifelse(nn5Dist_s > quantile(nn5Dist_s, 0.975), 1, 0)) %>%
    mutate(middle = ifelse(Ltail+Rtail == 0, 1, 0))
  
  redata <- reshape(test, varying = list(vars), direction="long", timevar="varb", 
                    times = as.factor(vars), v.names="resp")
  
  redata <- data.frame(redata, root = 1)
  redata$fit <- as.numeric(redata$varb == "embryoCt")
  redata$Nid <- as.numeric(gsub("[^0-9.-]", "", redata$id))
  redata$Deer <- as.numeric(redata$varb=='flCtNotConsumed')
  redata$Pollination <- as.numeric(is.element(redata$varb, 
                                c("capsuleCt", "isHarvested", "ovuleCt", "embryoCt")))

  
  return(redata)
}
```

## Zooming-in to the middle range of original landscapes

```{r}
#par(mar = c(4, 4, .1, .1))
fitness_landscape(model1, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model2, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model3, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model4, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model5, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model6, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model7, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model8, covariate = 'nn5DistNotConsumed_s', lower = quantile(test$nn5DistNotConsumed_s, 0.025), upper = quantile(test$nn5DistNotConsumed_s, 0.975))
```


```{r}
#par(mar = c(4, 4, .1, .1))
fitness_landscape(model1, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model2, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model3, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model4, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model5, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model6, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model7, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE)
fitness_landscape(model8, covariate = 'nn5DistNotConsumed_s', lower = quantile(test$nn5DistNotConsumed_s, 0.025), upper = quantile(test$nn5DistNotConsumed_s, 0.975),  observation = TRUE)
```

```{r}
#par(mar = c(4, 4, .1, .1))
fitness_landscape(model1, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model2, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model3, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model4, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model5, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model6, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model7, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975),  observation = TRUE,scale_back = TRUE)
fitness_landscape(model8, covariate = 'nn5DistNotConsumed_s', lower = quantile(test$nn5DistNotConsumed_s, 0.025), upper = quantile(test$nn5DistNotConsumed_s, 0.975),  observation = TRUE,scale_back = TRUE)
```


## Refit models on trimmed data 

```{r}
model_selection <- function(redata) {
  mod.null <- aster(resp ~ -1 + varb, pred, fam,varb,id,root,data=redata)
  mod1 <- aster(resp ~ -1 + varb + fit:nn5Dist_s, pred, fam,varb,id,root,data=redata)
  mod2 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s, pred, fam,varb,id,root,data=redata)
  mod3 <- aster(resp ~ -1 + varb + Pollination:nn5Dist_s, pred, fam,varb,id,root,data=redata)
  mod4 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s, pred, fam,varb,id,root,data=redata)
  mod5 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Pollination:nn5Dist_s, pred, fam,varb,id,root,data=redata)
  mod6 <- aster(resp ~ -1 + varb + Deer:nn5Dist_s + Pollination:nn5Dist_s, pred, fam,varb,id,root,data=redata)
  mod7 <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s, 
                  pred, fam,varb,id,root,data=redata)
  mod8 <- aster(resp ~ -1 + varb + Pollination:nn5DistNotConsumed_s, pred, fam,varb,id,root,data=redata)
  AICs <- rep(0, 9)
  AICs[1] <- mod.null$deviance + 2*length(mod.null$coefficients)
  AICs[2] <- mod1$deviance + 2*length(mod1$coefficients)
  AICs[3] <- mod2$deviance + 2*length(mod2$coefficients)
  AICs[4] <- mod3$deviance + 2*length(mod3$coefficients)
  AICs[5] <- mod4$deviance + 2*length(mod4$coefficients)
  AICs[6] <- mod5$deviance + 2*length(mod5$coefficients)
  AICs[7] <- mod6$deviance + 2*length(mod6$coefficients)
  AICs[8] <- mod7$deviance + 2*length(mod7$coefficients)
  AICs[9] <- mod8$deviance + 2*length(mod8$coefficients)
  print(which.min(AICs))
}
```

```{r}
min10 <- sort(data$nn5Dist)[1:10]
max10 <- sort(data$nn5Dist, decreasing = TRUE)[1:10]
```


```{r}
data_copy <- data


redata_trimmed <- wide2long(data_copy %>% filter(nn5Dist > quantile(nn5Dist, 0.0125)))
mod <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s,
              pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape(mod)

redata_trimmed <- wide2long(data_copy %>% filter(nn5Dist < quantile(nn5Dist, 0.9875)))
mod <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s,
              pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape(mod)

redata_trimmed <- wide2long(data_copy %>% filter(nn5Dist > quantile(nn5Dist, 0.0125), 
                                                  nn5Dist < quantile(nn5Dist, 0.9875)))
mod <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s,
              pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape(mod)

redata_trimmed <- wide2long(data_copy %>% filter(nn5Dist > quantile(nn5Dist, 0.025)))
mod <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s,
              pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape(mod)

redata_trimmed <- wide2long(data_copy %>% filter(nn5Dist < quantile(nn5Dist, 0.975)))
mod <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s,
              pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape(mod)

redata_trimmed <- wide2long(data_copy %>% filter(nn5Dist > quantile(nn5Dist, 0.025), 
                                                  nn5Dist < quantile(nn5Dist, 0.975)))
mod <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s,
              pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape(mod)


```

```{r}
aster_AIC(mod)
```


```{r}
mod_2d <- aster(resp ~ -1 + varb + fit:nn5Dist_s +Deer:nn5Dist_s + Pollination:nn5DistNotConsumed_s, 
                pred, fam,varb,id,root,data=redata_trimmed)
fitness_landscape_2d(mod_2d)
```

```{r}
fitness_landscape_2d(model11)
```

## Modelling tail behavior with higher order



```{r}
aster_AIC(model7)
```

### Quadratic and cubic



```{r}
model_comp <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s + 
                      fit:I(nn5Dist_s^2) + Deer:I(nn5Dist_s^2) + Pollination:I(nn5Dist_s^2), 
                pred, fam,varb,id,root,data=redata)

#summary(model_comp)
aster_AIC(model_comp)
#summary(model_comp, info.tol = 1e-9)
```

```{r}
model_cubic1 <- aster(resp ~ -1 + varb + fit:(nn5Dist_s+I(nn5Dist_s^3)) + Deer:(I(nn5Dist_s^2)) +
                      Pollination:(nn5Dist_s+I(nn5Dist_s^3) + I(nn5Dist_s^2)) , 
                pred, fam,varb,id,root,data=redata)

#summary(model_comp)
aster_AIC(model_cubic1)
summary(model_cubic1, info.tol=1e-8)
```

```{r}
fitness_landscape(model_cubic1, observation = TRUE)
fitness_landscape(model_cubic1, observation = TRUE, scale_back = TRUE)
fitness_landscape(model_cubic1, lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975))
fitness_landscape(model_cubic1,lower = quantile(log(data$nn5Dist)/10, 0.025), upper = quantile(log(data$nn5Dist)/10, 0.975), observation = TRUE, scale_back = TRUE)
```
### Piecewise linear

```{r}
model_comp <- aster(resp ~ -1 + varb + fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s + 
                      Ltail:nn5Dist_s + Rtail:nn5Dist_s, 
                pred, fam,varb,id,root,data=wide2long(data_copy))

#summary(model_comp)
aster_AIC(model_comp)
#summary(model_comp, info.tol=1e-8)
```

```{r}
model_comp <- aster(resp ~ -1 + varb + middle:(fit:nn5Dist_s + Deer:nn5Dist_s + Pollination:nn5Dist_s) + 
                      Ltail:(fit:nn5Dist_s) + 
                      Rtail:(fit:nn5Dist_s  + Pollination:nn5Dist_s), 
                pred, fam,varb,id,root,data=wide2long(data_copy))

#summary(model_comp)
aster_AIC(model_comp)
summary(model_comp, info.tol=1e-11)
```

### Piecewise quadratic/cubic

```{r}
model_comp <- aster(resp ~ -1 + varb  +
                      Rtail:(fit:(I(nn5Dist_s^2)) + Deer:(nn5Dist_s) +Pollination:( I(nn5Dist_s^2))) +
                      fit:(nn5Dist_s+I(nn5Dist_s^2) ) + Deer:(I(nn5Dist_s^3)) +
                      Pollination:(nn5Dist_s + I(nn5Dist_s^2)+ I(nn5Dist_s^3))+
                      Ltail:(fit:(nn5Dist_s) + Pollination:(nn5Dist_s + I(nn5Dist_s^2))), 
                pred, fam,varb,id,root,data=wide2long(data_copy))

#summary(model_comp)
aster_AIC(model_comp)
summary(model_comp, info.tol=1e-12)
```




## Third order model on trimmed data 

```{r}
model_cubic2 <- aster(resp ~ -1 + varb + fit:(nn5Dist_s+I(nn5Dist_s^2)) + 
                      Deer:(I(nn5Dist_s^2)) +
                      Pollination:(nn5Dist_s+I(nn5Dist_s^2) + I(nn5Dist_s^3)) , 
                pred, fam,varb,id,root,data=redata_trimmed)

#summary(model_comp)
aster_AIC(model_cubic2)
summary(model_cubic2, info.tol=1e-8)

```

```{r}
fitness_landscape(model_cubic2)
fitness_landscape(model_cubic2, observation = TRUE)
fitness_landscape(model_cubic2, scale_back = TRUE)
fitness_landscape(model_cubic2, observation = TRUE, scale_back = TRUE)
```


# Conditional landscapes

According to Jared and Prof. Wagenius, the U-shape landscape is somehow against biological intuitive. Let's see conditonal landscape of each node.


```{r}
conditional_fitness_landscape <- function(model, covariate = 'nn5Dist_s', lower = NULL, upper = NULL, 
                              observation=FALSE, scale_back = FALSE) {
  # Make fake individuals
  nInd <- 50
  lwr <- if (is.null(lower)) min(unique(model$data[,covariate])) else lower
  upr <- if (is.null(upper)) max(unique(model$data[,covariate])) else upper
  cand.nnA <- seq(from = lwr, to = upr, length = nInd)
  cand <- as.data.frame(cand.nnA)
  colnames(cand) <- covariate
  cand$root <- 1
  blah <- data[1:nInd, colnames(data) %in% vars]
  cand <- cbind(cand, blah)
  cand$id <- data[1:nInd, 'id']
  
  # Transform fake data into long format
  cand_long <- reshape(cand, varying = list(vars), direction="long", timevar="varb", 
                  times = as.factor(vars), v.names="resp")

  cand_long <- data.frame(cand_long)
  cand_long$fit <- as.numeric(cand_long$varb == "embryoCt")
  cand_long$Nid <- as.numeric(gsub("[^0-9.-]", "", cand_long$id))
  
  cand_long$Deer <- as.numeric(cand_long$varb=='flCtNotConsumed') 
  cand_long$Pollination <- as.numeric(is.element(cand_long$varb, 
                                c("capsuleCt", "isHarvested", "ovuleCt", "embryoCt")))
  cand_long$Rtail <- as.numeric(cand_long$nn5Dist_s > quantile(cand_long$nn5Dist_s, 0.975)) 
  cand_long$Ltail <- as.numeric(cand_long$nn5Dist_s < quantile(cand_long$nn5Dist_s, 0.025)) 
  
  # Get conditional mean value parameters
  pred <- predict(model, cand_long, varvar=varb, idvar=id, root=root,
                      se.fit = TRUE, model.type='conditional',
                  is.always.parameter = TRUE, info.tol=1e-8)
  
  xi_parm <- pred$fit
  xi_parm_se <- pred$se.fit
  
  names(xi_parm) <- paste0(cand_long$id,'.',cand_long$varb)
  names(xi_parm_se) <- paste0(cand_long$id,'.',cand_long$varb)
  
  conditional_exp <- matrix(xi_parm, nrow=50)
  rownames(conditional_exp) <- cand$id
  colnames(conditional_exp) <- paste0(vars,'.exp')
  
  conditional_exp_se <- matrix(xi_parm_se, nrow=50)
  rownames(conditional_exp_se) <- cand$id
  colnames(conditional_exp_se) <- paste0(vars,'.exp_se')
  
  
  cand_block <- cbind(cand, conditional_exp, conditional_exp_se)
  
  xlabel = if (covariate == 'nn5Dist_s') 'log(nnA)/10' else 'log(nnB)/10'

  #cand_block <- cand %>% mutate(xi_parm[1:50], xi_parm, lower = xi_parm[1:50] - 2 * xi_parm_se[1:50], upper = xi_parm[1:50] + 2 * xi_parm_se[1:50])
  if (scale_back == TRUE) {
    cand_block[, as.character(covariate)] <- exp(10*cand_block[,as.character(covariate)])
    xlabel = if (covariate == 'nn5Dist_s') 'nnA' else 'nnB'
  }
  #cand_block <- cand_block[1:40,]
  for (node in vars) {
    plt <- ggplot(data = cand_block) + geom_line(mapping = aes(x = cand_block[,as.character(covariate)], y = cand_block[, paste0(node, '.exp')])) + geom_ribbon(mapping = aes(x = cand_block[,as.character(covariate)], ymin = cand_block[, paste0(node, '.exp')] - 2*cand_block[, paste0(node, '.exp_se')], ymax = cand_block[, paste0(node, '.exp')] + 2*cand_block[, paste0(node, '.exp_se')]), alpha = 0.5)
    if (observation == TRUE) {
      obs <- test %>% filter((!!sym(covariate)) > lwr & (!!sym(covariate)) < upr)
      if (scale_back == TRUE) {
        plt <- plt + geom_point(data=obs, mapping = aes(x =exp(10*obs[,as.character(covariate)]),
                                                        y = obs[,'fecundity']))
      } else {
        plt <- plt + geom_point(data=obs, mapping = aes(x =obs[,as.character(covariate)], y = obs[,'fecundity']))
      }
    }
    plt <- plt + labs(x=xlabel, y="Expected fitness", title = paste0("Conditional fitness landscape of   ", node))  +
      #annotate(geom='text', x=c(0, 0.1, 0.3, 0.5), y=750, label=c('nnA', '2.718282', '20.085537', '148.413159')) +
      theme_minimal() + scale_x_continuous(minor_breaks = NULL) +scale_y_continuous(minor_breaks = NULL)
  
    print(plt)
  }
  
  
}
```


```{r}
conditional_fitness_landscape(model7)
```

```{r}
conditional_fitness_landscape(model_cubic1)
```

```{r}
conditional_fitness_landscape(model_cubic2)
```


# Zero-inflated model

As a comparison, a zero-inflated poisson model is fitted.

```{r}
library(pscl)
```

```{r}
mod_zeroinfl <- zeroinfl(embryoCt ~ nn5Dist_s, dist = 'poisson', data = test)
summary(mod_zeroinfl)
```

```{r}
landscape_zeroinfl <- function(model, covariate = 'nn5Dist_s', lower = NULL, upper = NULL, 
                              observation=FALSE, scale_back = FALSE) {
  # Make fake individuals
  nInd <- 50
  lwr <- if (is.null(lower)) min(unique(model$model[,covariate])) else lower
  upr <- if (is.null(upper)) max(unique(model$model[,covariate])) else upper
  cand.nnA <- seq(from = lwr, to = upr, length = nInd)
  cand <- as.data.frame(cand.nnA)
  colnames(cand) <- covariate
  cand$root <- 1
  blah <- data[1:nInd, colnames(data) %in% vars]
  cand <- cbind(cand, blah)
  cand$id <- data[1:nInd, 'id']
  
  #print(head(cand))
  
  
  # Get conditional mean value parameters
  pred <- predict(model, newdata=cand, type='response')
  #print(pred)
  
  pred_prob <- predict(model, newdata=cand, type='prob')
  #print(dim(pred_prob))
  
  xlabel = if (covariate == 'nn5Dist_s') 'log(nnA)/10' else 'log(nnB)/10'
  if (scale_back == TRUE) {
    cand[, as.character(covariate)] <- exp(10*cand[,as.character(covariate)])
    xlabel = if (covariate == 'nn5Dist_s') 'nnA' else 'nnB'
  }
  #cand_block <- cand_block[1:40,]
  plt <- ggplot(data = cand) + geom_line(mapping = aes(x = cand[,as.character(covariate)], y = pred))
  if (observation == TRUE) {
    obs <- test %>% filter((!!sym(covariate)) > lwr & (!!sym(covariate)) < upr)
    if (scale_back == TRUE) {
      plt <- plt + geom_point(data=obs, mapping = aes(x =exp(10*obs[,as.character(covariate)]), 
                                                      y = obs[,'fecundity']))
    } else {
      plt <- plt + geom_point(data=obs, mapping = aes(x =obs[,as.character(covariate)], y = obs[,'fecundity']))
    }
  }
  plt <- plt + labs(x=xlabel, y="Expected fitness", title = paste0("Fitness landscape    ", substr(format(model$formula), start=20, stop=nchar(format(model$formula)))))  + 
    #annotate(geom='text', x=c(0, 0.1, 0.3, 0.5), y=750, label=c('nnA', '2.718282', '20.085537', '148.413159')) +
    theme_minimal() + scale_x_continuous(minor_breaks = NULL) +scale_y_continuous(minor_breaks = NULL)
  
  print(plt)
  return(pred_prob)
}
```

```{r}
zeroinfl_pred_prob <- landscape_zeroinfl(mod_zeroinfl)
```


