---
title: "Investigation on sub-sampling"
author: "Gan Yao"
date: "`r Sys.Date()`"
output: pdf_document
---

Load libraries and read data.

```{r message=FALSE}
library(aster)
library(tidyverse)
redata <- read.csv("Lilium_processed.csv")
redata$site <- as.factor(redata$site)
redata$varb <- as.factor(redata$varb)
```

First fit the full random effects model.

```{r}
pred <- c(0,1,2,3,4,5)
fam <- c(2,1,2,1,2,1)
redata$Nid <- as.numeric(gsub("[^0-9.-]", "", redata$id))
remodel <- reaster(resp ~ varb + fit:(nn1Dist + nn2Dist + nn3Dist + nn4Dist + nn5Dist
                                       + nn6Dist + nn7Dist + nn8Dist + nn9Dist + nn10Dist), 
                    list(site = ~0 + fit:site),
                    pred, fam,varb,Nid,root,data=redata)
```

And a fixed effects model to help us with the transformation.

```{r}
femodel <- aster(resp ~ varb + fit:(site+nn1Dist + nn2Dist + nn3Dist + nn4Dist + nn5Dist
                                       + nn6Dist + nn7Dist + nn8Dist + nn9Dist + nn10Dist), 
                    pred, fam,varb,Nid,root,data=redata)
```

Get our "new data", which is just data for one individual.

```{r eval=FALSE, include=FALSE}
id <- redata$id
inies <- id == min(id)
#inies <- id == 'LP1012'
newdata <- redata[inies, ]
dim(newdata)
```

Get the parameters of random effects model for prediction.

```{r}
alpha.hat <- remodel$alpha
b.hat <- remodel$b
fred <- c(alpha.hat, b.hat)
idx <- match(names(femodel$coefficients), names(fred))
idx
fred[-idx]
```

Now we do the prediction.

```{r error=TRUE}
pred <- predict(femodel, varvar=varb, idvar=id, root=root, se.fit=TRUE, newcoef=fred[idx], model.type = 'conditional')
pred <- predict(femodel, varvar=varb, idvar=id, root=root, se.fit=TRUE, newdata=newdata, newcoef=fred[idx], model.type = 'conditional')
```

Direction of recession problem pops out again. Set `info.tol`.

```{r}

#Prediction for all individuals, with `is.always.parameter` being TRUE
pred <- predict(femodel, varvar=varb, idvar=id, root=root, se.fit=TRUE, newcoef=fred[idx], model.type = 'conditional', info.tol  = 1e-11, is.always.parameter = TRUE)
#uncond_pred <- predict(femodel, varvar=varb, idvar=id, root=root, se.fit=TRUE, newcoef=fred[idx], info.tol  = 1e-11, is.always.parameter = TRUE)


xis <- pred$fit
names(xis) <- redata$X

```

Get variation of mean value parameters by delta method.

```{r}
foo <- pred$gradient
rownames(foo) <- redata$X
colnames(foo) <- names(femodel$coefficients)
thegradient <- foo[, "fit:sitebeng"]
thevariance <- thegradient %*% t(thegradient) * remodel$nu
rownames(thevariance) <- redata$X
```

Expected fitness should be $\xi_1\xi_2\xi_3\xi_5\xi_6$.

```{r}
exp_fitness <- rep(0,697)
for (i in 1:697) {
  ind <- orgdata[i, 'id']
  exp_fitness[i] <- prod(xis[grep(ind, names(xis))][-4])
}
```

Get variation of expected fitness by delta method.

```{r}
var_expfit <- rep(0,697)
for (i in 1:697) {
  ind <- orgdata[i, 'id']
  ind <- paste(ind,'\\.', sep='')
  xi <- xis[grep(ind, names(xis))][-4]
  var <- thevariance[grep(ind, rownames(thevariance))[-4], grep(ind, colnames(thevariance))[-4]]
  grad <- c(prod(xi[-1]), prod(xi[-2]), prod(xi[-3]), prod(xi[-4]), prod(xi[-5]))
  var_expfit[i] <- t(grad) %*% var %*% grad
}
```




```{r eval=FALSE, include=FALSE}
orgdata <- read.csv("data/output/remLilium2021Data.csv")
orgdata <- orgdata[orgdata$site != "lf",]
orgdata <- orgdata[orgdata$site != "wrrx",]
orgdata[is.na(orgdata$nCapsulesHarvested), 'nCapsulesHarvested'] <- 0
orgdata %>% group_by(site) %>% summarize(mean(nCapsulesHarvested))
```

```{r eval=FALSE, include=FALSE}
test.pred <- predict(femodel, varvar=varb, idvar=id, root=root,  se.fit=TRUE, newcoef=fred[idx], model.type="conditional", info.tol=1e-11)


cond.pred <- matrix(test.pred$fit, ncol=6)
colnames(cond.pred) <- c('flCt_c','flNotCon_c','capsuleCt_c','isHarvested_c','ovuleCt_c','embryoCt_c')
org_b_cond <- cbind(orgdata, cond.pred)

org_b_cond %>% group_by(site, flCt) %>% summarize(mean(nCapsulesHarvested))

test.se_mat <- matrix(test.pred$se.fit, ncol=6)
```
